Put Title here

### 2018-12-11 General WLD combinatorics.tex
various stuff I was supposed to do in section 3

Start by choosing a propagator $p_1$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+2)$ for its support.  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise: 
Since $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex is supported then there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  This bounds a region $[i_1,k_1]$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

@@

\begin{lem}\label{lem last vert}
  Let W be an admissible WLD on $n$ vertices with Grassmann necklace $I_1, \ldots, I_n$.  Then for all $1\leq a \leq n$, $a-1 \not\in I_a$ interpreted cyclically (that is for $a=1$, $a-1$ is interpreted to be $n$.)
\end{lem}

\begin{proof}
Without loss of generality we can suppose $a=1$, so it suffices to show that $C(W)$ with the $n$th column removed is full rank.  (This suffices because if $C(W)$ with the $n$th column removed had full rank but the lexicographically minimal basis of $C(W)$ included the $n$th column, then we'd obtain a contradiction by taking this minimal basis with the $n$th column removed and extending it to a basis using a column from the first $n-1$ columns of $C(W)$ which is possible by the full rank assumption.)

Because the nonzero entries of $C(W)$ are independent indeterminants, it suffices to show that there is some choice of nonzero entries of $C(W)$, one in each row, each in different columns, and none of which are the $n$th column.  This is equivalent to an assignment of all propagators to supporting vertices so that no two are assigned to the same vertex and none are assigned to vertex $n$.  The lemma is proved if we show such an assignment exists.

Consider an assignment $f$ of a subset of propagators to vertices not including vertex $n$ with the maximal number of propagators.  If $f$ assigns $n$ propagators then we are done.  Otherwise there is a propagator $p$ which is not assigned but, by maximality, all of $p$s supporting vertices (other than possibly $n$) are assigned.  Build a sequence of sets $S_i$ of propagators and $T_i$ of vertices as follows.

\begin{enumerate}
  \item $S_0={p}$ and $T_0$ is the support of $p$ except for $n$.  Set $i=0$.
  \item Let $R$ be the set of propagators which are assigned to a vertex in $T_i$ by $f$. Let $S_{i+1}= S_i \cup R$.  Let $T_{i+1} = T_i \cup (\text{supp}(R) - \{n\})$.\label{item loop step}
  \item If there is a vertex $v_{i+1}$ in $T_{i+1}$ which has not been assigned by $f$ then by construction there is a propagator $q_{i+1} \in S_{i+1}-S_i$ which can be assigned to $v_{i+1}$.  Do so, removing $q_{i+1}$'s old assignment.  This frees up a vertex $v_{i}$ in $T_i$ and so there is a propagator $q_{i} \in S_{i}$ which can be assigned to this vertex freeing up an earlier vertex and so on until $p$ has been assigned.  This new assignment contradicts the maximality of $f$.
  \item  If nothing changed in step \ref{item loop step} then all vertices of $T_{i+1}$ are assigned to propagators in $S_{i+1}-\{p\}$ and vice versa.  Thus $|S_{i+1}| = |T_{i+1}|+1$.  Also all the support of propagators in $S_{i+1}$ is in $T_{i+1} \cup \{n\}$, which is a set of size $|T_{i+1}|+1 = |S_{i+1}|$.  Thus we have $|S_{i+1}|$ propagators supported on $|S_{i+1}|$ vertices contradicting admissibility.
  \item Increment $i$ and go to \ref{item loop step}
\end{enumerate}

This procedure must be finite as either a contradiction is found or $S_i$ grows at each step but $W$ has finitely many propagators.  However, the only way for the process to end is by reaching a contradiction.  Therefore any maximal assignment is a basis not including column $n$ which proves the lemma.
\end{proof}

%such that $b+1 \in V_{p_1}$, with an endpoint more clockwise on $W$ than the endpoint of $p$ supported by $b+1$: see Figure~\ref{fig:base step of alg induction}.
%
%\begin{figure}
%\begin{tikzpicture}[rotate=15]
%\begin{scope}
%\drawWLD{12}{2}
%\drawprop{3}{2}{10}{2} % P
%% number vertices a,a+1,b,b+1 for prop P
%
%% clip these
%\drawprop{3}{-2}{12}{-2}
%\drawprop{2}{0}{12}{2}
%\drawprop{6}{0}{10}{-2} % p1
%
%
%\end{scope}
%\end{tikzpicture}
%\caption{Base step: if $P$ survives $n$ steps of the algorithm, there must be another propagator $p_1$ to its right on edge $(b,b+1)$.}
%\label{fig:base step of alg induction}
%\end{figure}
%
%The propagator $p_1$ forms the base step for the following induction:
%
%Suppose that $p_r$ is a propagator supported on $(i,i+1,m,m+1)$ and removed at step $m+1$ of the algorithm; we will show that there must also be a propagator $p_{r+1}$ supported on $(i',i'+1,m',m'+1)$ with $i \leq i'$ and $m \geq m'$, which is removed at step $m'+1$.  Since we can't have both $i' = i$ and $m'=m$, $p_{r+1}$ bounds a region containing strictly fewer vertices than $p_r$.
%
%Notice that if we restrict our attention to the region bounded by $p_r$ and the edges of the diagram from $i$ to $m+1$, then propagators outside of this region can affect the algorithm only at vertices $i$, $i+1$.  (The starting vertex 1 lies outside this region, by assumption.)
%
%Given a propagator $p_r$ supported on $(i,i+1,m,m+1)$ and removed at step $m+1$, we must also have a propagator $q$ to its right which is removed at step $m$ (``protecting'' $p_r$ at step $m$).  This splits into two cases:
%
%\textbf{Case I}: If $q$ is supported on $(k,k+1,m-1,m)$ for some $k \geq i$, then $p_{r+1}:=q$  satisfies the conditions of the induction step.
%
%\textbf{Case II}: Suppose $q$ is supported on $(k,k+1,m,m+1)$ for some $k$; since the diagram is admissible, we have $k > i$.  In particular, anything that happens at vertex $k+1$ is unaffected by propagators outside the region bounded by $p_r$, since $k+1 > i+1$.  But $q$ must survive until step $m$, so there must be another propagator $q'$ to the right of $q$, which protects $q$ at step $k+1$.  The admissibility of $W$ implies that $q'$ is supported on $(j,j+1,k,k+1)$ for some $j \geq i$, and we can take $p_{r+1}:=q'$.
%
%\begin{figure}
%\begin{tabular}{ccc}
%\begin{tikzpicture}[rotate=15]
%\begin{scope}
%\drawWLD{12}{2}
%\drawprop{3}{0}{10}{0} % p_r
%% number vertices i,i+1,m,m+1
%
%%clip these
%\drawprop{6}{0}{9}{0} % q
%\end{scope}
%\end{tikzpicture}
%
%& \qquad \qquad &
%\begin{tikzpicture}[rotate=15]
%\begin{scope}
%\drawWLD{12}{2}
%\drawprop{3}{0}{10}{2} % p_r
%% number vertices i, i+1, m, m+1
%
%\drawprop{7}{2}{10}{-2} %q
%% vertices k,k+1, m,m+1
%
%% clip this one
%\drawprop{4}{0}{7}{-2} % q'
%% vertices j,j+1,k, k+1
%
%\end{scope}
%\end{tikzpicture}
%\\
%Case I & & Case II
%\end{tabular}
%
%\caption{Inductive step: if propagator $q$ is removed at step $m$, then it lies on either edge $(m-1,m)$ (Case I) or $(m,m+1)$ (Case II).}
%\label{fig: cases for inductive step, algorithm}
%\end{figure}
%
%By induction, we will eventually find a propagator on a region with 3 or fewer v%ertices, which is inadmissible.

\begin{lem} \label{vertex cyclic int lem}
  For every $v \in V_p$, Let $A_v \subset [n]$ be the set of vertices where $p$ contributes $v$ to $I_a$. Then $A_v$ is a non-empty cyclic interval.

  \note{the above is the statement that was in this file, the below is the version of the statement from last july}

Let $p$ be a propagator of an admissible WLD $W$ on $n$ vertices.  Let $p = (i,j)$.  For $l\in \{i, i+1, j, j+1\}$, let $J_p^{(W)}(l)$ be the indices $m$ where $I_m(p)=l$.  Then $J_p^{(W)}(i)$, $J_p^{(W)}(i+1)$, $J_p^{(W)}(j)$, and $J_p^{(W)}(j+1)$ are each non-empty cyclic intervals of $\{1, \ldots, n\}$ which partition $\{1, \ldots, n\}$ and occur in the given cyclic order.
\end{lem}

@@

Now we proceed to consider the first three cases of Figure~\ref{fig 3 cases}.
We first describe $J_q^{(V)}(*)$: since there are no propagators inside $q$, we see from the diagram that
\[l, l+1, \ldots ,j \in J^{(V)}_q(j) \text{ (if $l$ exists) and }  j+1 \in J^{(V)}_q(j+1).\]
Note that $j+2 \not\in J^{(V)}_q(j+1)$ by Lemma~\ref{lem last vert}, so $J^{(V)}_q(j+1) = \{j+1\}$ and $j+2 \in J^{(V)}_q(i)$ by the induction hypothesis. Thus there exist vertices $d,e \in [j+2,i+1]$ with $d <e$, such that
\[J^{(V)}_q(i) = [j+2,d-1], \quad J^{(V)}_q(i+1) = [d,e-1], \quad J^{(V)}_q(j) = [e,j], \quad J^{(V)}_q(j+1) = \{j+1\},\]
and all intervals are non-empty.

@@

>>>>

Start by choosing a propagator $p_1$ of length 2 in $W$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+2)$ for its support.  If it is part of one of the configurations we are looking for then we are done, 
%a configuration \eqref{eq:2 prop config} we are done,
so suppose otherwise: 
By assumption $p_1$ is not in configuration \ref{item pair of 2s}, there are no propagators of length $3$, and every vertex is supported.  Therefore there must exist a propagator $q_1$ of length $\geq 4$ with one end on edge $j_1$ or on edge $j_1-1$ or on edge $j_1-2$.  This bounds a region $[i_1,k_1]$ (where $k_1 \in \{j_1-1, j_1, j_1+1\}$) which does not contain $p_1$.

@@

%\begin{lem}\label{lem last vert}
%  Let W be an admissible WLD on $n$ vertices with Grassmann necklace $I_1, \ldots, I_n$.  Then for all $1\leq a \leq n$, $a-1 \not\in I_a$ interpreted cyclically (that is for $a=1$, $a-1$ is interpreted to be $n$.)
%\end{lem}
%
%\begin{proof}
%Without loss of generality we can suppose $a=1$, so it suffices to show that $C(W)$ with the $n$th column removed is full rank.  (This suffices because if $C(W)$ with the $n$th column removed had full rank but the lexicographically minimal basis of $C(W)$ included the $n$th column, then we'd obtain a contradiction by taking this minimal basis with the $n$th column removed and extending it to a basis using a column from the first $n-1$ columns of $C(W)$ which is possible by the full rank assumption.)
%
%Because the nonzero entries of $C(W)$ are independent indeterminants, it suffices to show that there is some choice of nonzero entries of $C(W)$, one in each row, each in different columns, and none of which are the $n$th column.  This is equivalent to an assignment of all propagators to supporting vertices so that no two are assigned to the same vertex and none are assigned to vertex $n$.  The lemma is proved if we show such an assignment exists.
%
%Consider an assignment $f$ of a subset of propagators to vertices not including vertex $n$ with the maximal number of propagators.  If $f$ assigns $n$ propagators then we are done.  Otherwise there is a propagator $p$ which is not assigned but, by maximality, all of $p$s supporting vertices (other than possibly $n$) are assigned.  Build a sequence of sets $S_i$ of propagators and $T_i$ of vertices as follows.
%
%\begin{enumerate}
%  \item $S_0={p}$ and $T_0$ is the support of $p$ except for $n$.  Set $i=0$.
%  \item Let $R$ be the set of propagators which are assigned to a vertex in $T_i$ by $f$. Let $S_{i+1}= S_i \cup R$.  Let $T_{i+1} = T_i \cup (\text{supp}(R) - \{n\})$.\label{item loop step}
%  \item If there is a vertex $v_{i+1}$ in $T_{i+1}$ which has not been assigned by $f$ then by construction there is a propagator $q_{i+1} \in S_{i+1}-S_i$ which can be assigned to $v_{i+1}$.  Do so, removing $q_{i+1}$'s old assignment.  This frees up a vertex $v_{i}$ in $T_i$ and so there is a propagator $q_{i} \in S_{i}$ which can be assigned to this vertex freeing up an earlier vertex and so on until $p$ has been assigned.  This new assignment contradicts the maximality of $f$.
%  \item  If nothing changed in step \ref{item loop step} then all vertices of $T_{i+1}$ are assigned to propagators in $S_{i+1}-\{p\}$ and vice versa.  Thus $|S_{i+1}| = |T_{i+1}|+1$.  Also all the support of propagators in $S_{i+1}$ is in $T_{i+1} \cup \{n\}$, which is a set of size $|T_{i+1}|+1 = |S_{i+1}|$.  Thus we have $|S_{i+1}|$ propagators supported on $|S_{i+1}|$ vertices contradicting admissibility.
%  \item Increment $i$ and go to \ref{item loop step}
%\end{enumerate}
%
%This procedure must be finite as either a contradiction is found or $S_i$ grows at each step but $W$ has finitely many propagators.  However, the only way for the process to end is by reaching a contradiction.  Therefore any maximal assignment is a basis not including column $n$ which proves the lemma.
%\end{proof}

Given a propagator $p$ of an admissible WLD $W$ and a vertex $i$ in the support of $p$ we will use the notation $J_p^{(W)}(i)$ for the set of indices $m$ where $I_m(p) = l$ where the $I_m$ are the sets calculated by Algorithm~\ref{alg:put GN on WLD}, and using the notation of Definition~\ref{def I_i as a function}.

\begin{lem} \label{vertex cyclic int lem}

Let $p=(i,j)$ be a propagator of an admissible WLD $W$ on $n$ vertices.  Then $J_p^{(W)}(i)$, $J_p^{(W)}(i+1)$, $J_p^{(W)}(j)$, and $J_p^{(W)}(j+1)$ are each non-empty cyclic intervals of $\{1, \ldots, n\}$ which partition $\{1, \ldots, n\}$ and occur in the given cyclic order.
\end{lem}

@@

Now we proceed to consider the first three cases of Figure~\ref{fig 3 cases}.
We first describe $J_q^{(V)}(*)$: since there are no propagators inside $q$, we see from the diagram that
\[l, l+1, \ldots ,j \in J^{(V)}_q(j) \text{ (if $l$ exists) and }  j+1 \in J^{(V)}_q(j+1).\]
Note that $j+2 \not\in J^{(V)}_q(j+1)$ by Lemma~\ref{lem no fourth vertex}, so $J^{(V)}_q(j+1) = \{j+1\}$ and $j+2 \in J^{(V)}_q(i)$ by the induction hypothesis. Thus there exist vertices $d,e \in [j+2,i+1]$ with $d <e$, such that
\[J^{(V)}_q(i) = [j+2,d-1], \quad J^{(V)}_q(i+1) = [d,e-1], \quad J^{(V)}_q(j) = [e,j], \quad J^{(V)}_q(j+1) = \{j+1\},\]
and all intervals are non-empty.

@@

We need one more lemma before we can prove that Algorithm~\ref{alg:put GN on WLD} does in fact give the Grassmann necklace of the positroid associated to $W$.

\begin{lem}\label{lem basis as perm}
Let $W = (\cP,[n])$ be an admissible Wilson loop diagram and let $M(W)$ be its associated matroid. A subset $J \subseteq [n]$ is an independent set of $M(W)$ if and only if there exists an injective set map $f : J \rightarrow \cP$ with the property that for each $s\in S$ we have $s \in V(f(s))$.
\end{lem}

One of the most important uses of this lemma is for bases.  The lemma says that a subset $J$ of $[n]$ is a basis of $M(W)$ if and only if there is a set bijection between $J$ and $\cP$ with the property that for each $j\in J$ the propagator associated to $j$ under the bijection is supported on vertex $j$.

\begin{proof}
  Because the nonzero entries of $C(W)$ are independent indeterminants,
  $J$ is an independent set is if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns.

  Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.

  Consequently, a choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statment, hence proving the result.
\end{proof}
